Technical Report

### 2018-03-21 fully-abstract-comp/technical-report.tex
Define linking, programs, contexts.

%\lau{16-10-2017: Fill this paragraph with $\scall{\offpc,\offsigma}{}{}$ implementation, i.e., the instructions $\scall[1]{\offpc,\offsigma}{r_1}{r_2} \dots \scall[\calllen]{\offpc,\offsigma}{r_1}{r_2}$ is equal to. Add convenience notation like $\update{\mem.a}{\scall{\offpc,\offsigma}{r_1}{r_2}} = \dots$}
\lau{16-10-2017: Implementation of $\scall{\offpc,\offsigma}{}{}$ may need registers for temporary registers - we should update the semantics accordingly.}
\[
  \begin{array}{l}
    \text{// push 42 on the stack (so it is non-empty).}\\
    \tmove{\rtmp{1}}{42}\\
    \tstore{\rstk}{\rtmp{1}}\\
    \tcca{\rstk}{(-1)}\\
    \text{// split the stack at its current address - rstk done.}\\
    \tgeta{\rtmp{1}}{\rstk}\\
    \tsplit{\rstk}{\rretd}{\rstk}{\rtmp{1}}\\
    \text{// load the seal for the return pointer through the pc capability.}\\
    \tmove{\rtmp{1}}{\pcreg}\\
    \tcca{\rtmp{1}}{(\offpc - 5)} \text{ // $\offpc$ is the offset from pc to the location of the seal for this code. \footnote{-5 is the offset from the previous instruction where the $\pcreg$ was moved to $\rtmp{1}$ to the first instruction of the call implementation.}}\\
    \tload{\rtmp{1}}{\rtmp{1}}\\

    \tcca{\rtmp{1}}{\offsigma}\text{ // $\offsigma$ is the offset for the seal used by this call.}\\
    \text{// seal the used stack frame as the data part of the return pointer pair.}\\
    \tcseal{\rretd}{\rtmp{1}}\\
    \text{// obtain the code part of the return pointer pair and seal it too.}\\
    \tmove{\rretc}{\pcreg}\\
    \tcca{\rretc}{5} \text{ //magic number is offset to return code}\\
    \tcseal{\rretc}{\rtmp{1}}\\
    \text{// now clear temporary register and jump to the adversary.}\\
    \tmove{\rtmp{1}}{0}\\
    \txjmp{r_1}{r_2}\\
    \text{// the following is the return code}\\
    \text{// check that the stack pointer is the same we handed out.}\\
    \tgetb{\rtmp{1}}{\rstk}\\
    \tminus{\rtmp{1}}{t_{t1}}{\stkb} \text{ //$\stkb$ is the stack base constant} \\
    \tmove{\rtmp{2}}{\pcreg}\\
    \tcca{\rtmp{2}}{5} \text{ //magic number is the offset to fail} \\
    \tjnz{\rtmp{2}}{\rtmp{1}} \\
    \tcca{\rtmp{2}}{1} \text{ //magic number is the offset to after fail} \\
    \tjmp{\rtmp{2}} \\
    \tfail \\
    \text{// join our stored private stack frame with the rest of the stack (this also finishes the stack pointer check).}\\
    \tsplice{\rstk}{\rstk}{\rdata} \\
    \text{// pop the magic number 42}\\
    \tcca{\rstk}{1}\\
    \text{// clear temporary registers used}\\
    \tmove{\rtmp{2}}{0}\\
    \text{// continue program after invocation.}
  \end{array}
\]
\dominique{Return code could be simplified if we had jz in addition to jnz.}
% \dominique{It is a bit annoying that we store the stack end address before invocation if we do not use it.
%   I understand this is for ensuring that our private stack frame is not empty, which is important.
%   But why don't we just push a constant then?
%   Also: perhaps we could only do it if the current private stack frame is empty to begin with?
% }
% \dominique{4-12-2017: update: Lau executed the first suggestion (push a constant instead), and convinced me that the last suggestion didn't make sense (only push if the current private stack is empty), because it is then not clear whether or not to pop in the return code.}
The call code does the following:
\begin{itemize}
\item Store 42 (it could be anything) to the stack (this ensures the stack is non-empty), and decrement the pointer according to convention.
\item Get the current address of the stack pointer and split according to it.
\item Retrieve the seal of the program.
\item cca the seal, so the seal to be used is active 
\item seal our private part of the stack capability.
\item Move the pc out of the pc register, adjust it to point to the first address of the return code, and seal it.
\item clear the temporary register.
\item cross jump to the two specified registers.
\item Upon return:
  \begin{itemize}
  \item get the base of the stack and check that it matches up with the global base of the stack.If not, fail.
  \item splice the returned stack pointer and the stack pointer for our private stack.
  \item adjust stack pointer to first empty address (the address with the end address is considered free)  
  \item clear the temporary registers. (Note that $\rtmp{1}$ is not cleared with 0 as it already contains 0 after the $\stkb$ check. If it didn't contain 0, then the execution would have failed.)

@@

TODO:
\begin{mathpar}
  \inference{
  }{
    \var{comp}_1 \bowtie \var{comp}_2 = \var{comp}_3
  }
\end{mathpar}

\subsection{Programs and contexts}
\begin{definition}
  \label{def:self-contained}
  \begin{itemize}
  \item We say that a memory segment $\ms$ and a list of words $\bar{w}$
    are self-contained iff
  \[
    \forall w \in \bar{w} \ldotp w \in \Caps \lor w = \stkptr{\_,\_,\_,\_} \Rightarrow \range{w} \subseteq \dom(\ms)
  \]
  and
  \begin{multline*}
    \forall a \in \dom(\ms) \ldotp\\
    (\ms(a) \in Caps \lor \ms(a) = \stkptr{\_,\_,\_,\_}) \Rightarrow \range{\ms(a)} \subseteq \dom(\ms)
  \end{multline*}
  \item We say that a memory segment $\ms$ and a list of capabilities $\bar{c}$ are self contained with slack $\ms_s$ iff
  \[
    \forall c \in \bar{c} \ldotp \range{c} \subseteq \dom(\ms \cup \ms_s)
  \]
  and
  \begin{multline*}
    \forall a \in \dom(\ms) \ldotp\\
    (\ms(a) \in Caps \lor \ms(a) = \stkptr{\_,\_,\_,\_}) \Rightarrow \range{\ms(a)} \subseteq \dom(\ms \cup \ms_s)
  \end{multline*}
\end{itemize}
\end{definition}

\begin{definition}
  \label{def:wxorx}
  ~\\
  A word $w$ is said to respect write-xor-execute iff
  \begin{itemize}
  \item $w \in \ints$
  \item $w = \seal{\_,\_,\_}$
  \item $w = ((\perm,\_),\_,\_,\_) \tand \perm \neq \rwx$
  \item $\sealed{\_,\vsc}$ and $\vsc$ respects write-xor-execute
  \item $\src{\stkptr{\perm} \tand \perm \neq \rwx}$
  \item $\src{w = \retptrc(\_)}$
  \item $\src{w = \retptrd(\_,\_)}$
  \end{itemize}
~\\
  A register file $\reg$ is said to respect write-xor-execute with slack $[\bar{r}]$ iff \\
  For all $r \in \RegName$ $\reg(r) \setminus [\bar{r}]$ respects write-xor-execute.
~\\
~\\
  A memory segment $\ms$ is said to respect write-xor-execute iff\\
  For all $\aaddr \in \dom(\ms)$ $\ms(\aaddr)$ respects write-xor-execute
\end{definition}

\begin{definition}
  \label{def:no-stk-ptr}
~\\
  We say that a word $w$ covers no stack pointer iff\\
  \begin{item}
    \item $w \neq ((\perm,\linear),\stkb,\_,\_)$\\
          $\src{w \neq \stkptr{\_,\_,\_,\_}}$
    \item if $w = \sealed{\_,\vsc}$, then $\vsc$ covers no stack pointer.
  \end{item}
~\\
  We say that a memory segment $\ms$ contains no stack pointer iff
  \[
    \forall \aaddr \in \dom(\ms)\ldotp \text{$\ms(\aaddr)$ covers no stack pointer}
  \]
~\\
  We say that a register file $\reg$ contains no stack pointer with slack $[\bar{r}]$ iff
  \[
    \forall r \in \reg \setminus [\bar{r}] \ldotp \text{$\reg(r)$ covers no stack pointer}
  \]
\end{definition}

\begin{definition}
  \label{def:resp-linearity}
~\\
  We say that a memory segment $\ms$ and words $\bar{w}$ respects linearity iff\\
  Let $\var{lc}$ be all the linear capabilities in $\ms$ and $\bar{w}$, i.e.,
  \[
    \var{lc} = \left\{c \middle|
      \begin{array}{l}
        \exists \aaddr \in \dom(\ms) \ldotp c = \ms(\aaddr) = ((\_,\linear),\_,\_,\_) \\
        \exists \aaddr \in \dom(\ms) \ldotp \ms(\aaddr) = \sealed{\_,c} \land c = ((\_,\linear),\_,\_,\_) \\
        \src{\exists \aaddr \in \dom(\ms) \ldotp c = \ms(\aaddr) = \stkptr{\_,\_,\_,\_}} \\
        \src{\exists \aaddr \in \dom(\ms) \ldotp \ms(\aaddr) = \sealed{\_,c} \land c= \stkptr{\_,\_,\_,\_}} \\
      \end{array}
\right\}
  \]
\end{definition}

\subsubsection{Programs}
Here we define what constitutes a program.
\lau{05-10-2017: It seems like both program and context takes care of linking atm.}
\begin{definition}
  \label{def:program}
  \lau{27-09-2017: The ``somehow''s need to be fixed.}
  A program, $\program$ consists of two sealed capabilities, $\vsc_\code = \sealed{\sigma,c_\code}$ and $\vsc_\data = \sealed{\sigma,c_\data}$, and two memory segments $\ms$ and $\ms_\link$ satisfying
  \begin{itemize}
  \item $c_\code = ((\rx,\lin_\code),\baddr_\code,\eaddr_\code,\aaddr_\code)$ where $\aaddr = \baddr + 2$ and $[\baddr_\code,\eaddr_\code]\subseteq \dom(\ms)$
  \item $c_\data = ((\rw,\lin_\data),\baddr_\data,\eaddr_\data,\baddr_\data)$ where $[\baddr_\data,\eaddr_\data] \subseteq \dom(\ms)$ and $[\baddr_\code,\eaddr_\code] \cap [\baddr_\data,\eaddr_\data] = \emptyset$
  \item $\ms$ and $[c_\code,c_\data]$ are self-contained with $\ms_\link$ as slack.
  \item $\ms$ respects write-xor-execute.
  \item $\ms(\baddr_\code) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\baddr}$ where these are somehow the sufficient amount of seals.
  \item $\ms(\baddr_\code+1) = ((\ro,\normal),\baddr_\link, \eaddr_\link, \baddr_\link)$ where $\dom(\ms_\link) = [\baddr_\link, \eaddr_\link]$
  \end{itemize}
\end{definition}
\lau{27-09-2017: At the moment, we do not have the proper support to let the code capability be linear (in the sense that you cannot write a sensible program) - maybe we should just require the code capability to be normal.}

\subsubsection{Contexts}
\begin{definition}
  \label{def:context}
  A context, $\context$, is a register file $\reg$, registers $r_1$ and $r_2$, and memory segments $\ms$ and $\ms_\link$ satisfying
  \begin{itemize}
  \item $\reg$ respects write-xor-execute with slack $[r_1,r_2]$.
  \item $\reg(\rstk) = ((\rw,\linear),\stkb,\eaddr,\eaddr)$ where $\stkb \leq \eaddr$\\
        {\sourcecolor{}$\reg(\rstk) = \stkptr{\rw,\stkb,\eaddr,\eaddr}$ where $\stkb \leq \eaddr$}
  \item $\ms \uplus \ms_\link$ respects write-xor-execute
  \item $\ms \uplus \ms_\link$ and $[\reg(\RegName)]$ are self-contained.
  \item $\ms \uplus \ms_\link$ contains no stack pointer.
  \item $\reg$ contains no stack pointer with slack $[r_1,r_2]$.
  \item {\color{red}TODO: No return pointer}
  \end{itemize}
In this definition, the source level specific conditions replace the equivalent target level conditions.
\end{definition}

\subsubsection{Plugging a context}

@@

\begin{definition}
  \label{def:points-to-call}
  A configuration {\sourcecolor $\Phi = (\ms,\reg,\_,\_)$} is said to point to instruction $\scall{o}{r_1}{r_2}$ in $A$ iff
  \[
    \reg(\pcreg) = ((\_,\_),\_,\eaddr,\aaddr) \tand [\aaddr,\aaddr+\calllen-1] \subseteq A
  \]
and
\[
  \aaddr+\calllen - 1 \leq \eaddr
\]
and
\[
  \callCond{\Phi,r_1,r_2,\aaddr}
\]
\end{definition}
We want to make a conditional full-abstraction statement. The condition should express two things (1) the compilation so far has been reasonable, so we do not have to setup protection against things any reasonable compiler wouldn't do. (2) the compilation so far has taken care of certain checks that we do not have information to place correctly at this point of the compilation.
\begin{definition}
  \label{def:check-stack-addr-before-call}
  A configuration $\src{\Phi}$ is said to check $\stkb$ before calls in $\ta$ for $n$ steps iff\\
  for all $i \leq n$ if
  \[
    \src{\Phi} \nstep[i] \src{\Phi'}
  \]
  and $\src{\Phi'}$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ in $\ta$ for some $\src{r_1}$ and $\src{r_2}$
  % and there exists no $j<i$ s.t.
  % \[
  %   \src{\Phi} \nstep[j] \src{\Phi''}
  % \]
  % where $\src{\Phi''}$ points to $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$ in $A$ for some $\src{r_1'}$ and $\src{r_2'}$,
  then
  \[
    \src{\Phi'}(\src{r_\stk}) = \src{\stkptr{\_,\stkb,\_,\_}}
  \]
  and
  \[
    r_1' \neq \rtmp{1}
  \]
  \lau{maybe the last bit should have its own definition.}
  % and $\src{\Phi'}$ has reasonable calls in $A$ for $n-i$ step.
\end{definition}

\begin{definition}[Use return seals only for call]
  \label{def:use-return-seals-call}
  A configuration $\src{\Phi}$ is said to use linker seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
  for all $i \leq n$, if
  \[
    \src{\Phi} \nstep[i] \src{\Phi'}
  \]
  and $\src{\Phi'}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi'(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, then either ($\src{\Phi'}$ is inside $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$ and $\sigma \in \sigrets$) or $\sigma \in \sigcloss$.
\end{definition}

\begin{definition}[Never pass trusted seals accross the trusted-untrusted boundary]
  \label{def:handle-trusted-seals-properly}
  A configuration $\src{\Phi}$ is said to handle trusted seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
  for all $i \leq n$, if
  \[
    \src{\Phi} \nstep[i] \src{\Phi'} \step \src{\Phi''}
  \]
  and $\src{\Phi'}$ is in $\ta$ and $\src{\Phi''}$ is not in $\ta$ and, then for all $r \in \RegName$
  %\dominique{TODO: $\sigma_b,\sigma_e$ contains a seal that was put in $\Phi$ by the linker.}
  \[
    \src{\Phi}.\reg(r) = \seal{\sigma_b,\sigma_e,\_} \Rightarrow (\sigrets \cup \sigcloss) \mathrel{\#} [\sigma_b,\sigma_e]
  \]
and
%  \dominique{TODO: Also not allowed: passing a capability to a code area that points (a.o.) to a memory location containing a seal like the one above.}
  \[
    \src{\Phi}.\reg(r) = ((\perm,\_),\baddr,\eaddr,\_) \wedge \perm \sqsubseteq \rx \Rightarrow [\baddr,\eaddr] \# \ta
  \]
  and
    \begin{multline*}
      \src{\Phi}.\reg(r) = \sealed{\sigma,\vsc} \wedge \sigma \not\in (\sigrets\cup\sigcloss) \Rightarrow 
      \\  \left(  
      \begin{array}{r l}
        (\vsc = \seal{\sigma_b,\sigma_e,\_} & \Rightarrow (\sigrets\cup\sigcloss) \# [\sigma_b,\sigma_e]) \wedge\\
        ( \vsc = ((\perm,\_),\baddr,\eaddr,\_) \wedge \perm \sqsubseteq \rx & \Rightarrow [\baddr,\eaddr] \# \ta )
      \end{array}
      \right)
  \end{multline*}
\end{definition}

%  \dominique{TODO: Also not allowed: storing a seal or a code capability in memory.}a
\begin{definition}[Never store link seals and code capability]
  \label{def:never-store-seal-code-cap}
  A configuration $\src{\Phi}$ is said to not store link seals and code capability in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
  for all $i \leq n$, if
  \[
    \src{\Phi} \nstep[i] \src{\Phi'}
  \]
  and $\src{\Phi'}$ points to $\src{\tload{r_1}{r_2}}$ in $\ta$, then 
\[
  \Phi'(r_2) = \seal{\sigma_\baddr,\sigma_\eaddr,\_} \Rightarrow (\sigrets\cup\sigcloss) \mathrel{\#} [\sigma_\baddr,\sigma_\eaddr]
\]
and
\[
  \Phi'(r_2) = ((\perm,\_),\baddr,\eaddr,) \wedge \perm \sqsubseteq \rx \Rightarrow [\baddr,\eaddr] \# \ta
\]
\end{definition}

@@

\dominique{19-3-2018: components should contain information about what seals they use and for what (for building return pointers or for building closures). That should facilitate defining reasonable linking (only when seals disjoint), and for defining the condition for full abstraction and the FTLR.}
\begin{definition}
  A component $\program = (\ms_{\mathrm{code}},\ms_{\mathrm{data}},\overline{c}_{\mathrm{import}},\overline{c}_{\mathrm{export}},\sigrets,\sigcloss)$ with $\dom(\ms_{\mathrm{code}}) = \ta$, is reasonable if for all $n$ and all contexts $\context$ the configuration $\plug{\context}{\program}$ satisfies:
  \begin{itemize}
  \item $\plug{\context}{\program}$ checks $\stkb$ before calls in $\ta$ for $n$ steps (Definition~\ref{def:check-stack-addr-before-call}).
  \item $\plug{\context}{\program}$ uses link seals properly in $\ta$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:use-return-seals-call}).
  \item $\plug{\context}{\program}$ handle trusted seals properly in $\ta$ with components parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:handle-trusted-seals-properly}).
  \item $\plug{\context}{\program}$ does not store link seals and code capabilities in $\ta$ with component parameters $\sigrets$ and $\sigcloss$ for $n$ steps (Definition~\ref{def:never-store-seal-code-cap}).
  \end{itemize}
\end{definition}

>>>>

%\lau{16-10-2017: Fill this paragraph with $\scall{\offpc,\offsigma}{}{}$ implementation, i.e., the instructions $\scall[1]{\offpc,\offsigma}{r_1}{r_2} \dots \scall[\calllen]{\offpc,\offsigma}{r_1}{r_2}$ is equal to. Add convenience notation like $\update{\mem.a}{\scall{\offpc,\offsigma}{r_1}{r_2}} = \dots$}
\lau{16-10-2017: Implementation of $\scall{\offpc,\offsigma}{}{}$ may need registers for temporary registers - we should update the semantics accordingly.}
\[
  \begin{array}{l}
    \text{// push 42 on the stack (so it is non-empty).}\\
    \tmove{\rtmp{1}}{42}\\
    \tstore{\rstk}{\rtmp{1}}\\
    \tcca{\rstk}{(-1)}\\
    \text{// split the stack at its current address - rstk done.}\\
    \tgeta{\rtmp{1}}{\rstk}\\
    \tsplit{\rstk}{\rretd}{\rstk}{\rtmp{1}}\\
    \text{// load the seal for the return pointer through the pc capability.}\\
    \tmove{\rtmp{1}}{\pcreg}\\
    \tcca{\rtmp{1}}{\offpc} \text{ // $\offpc$ is the offset from pc to the location of the seal for this code}\\
    \tload{\rtmp{1}}{\rtmp{1}}\\
    \tcca{\rtmp{1}}{\offsigma}\text{ // $\offsigma$ is the offset for the seal used by this call.}\\
    \text{// seal the used stack frame as the data part of the return pointer pair.}\\
    \tcseal{\rretd}{\rtmp{1}}\\
    \text{// obtain the code part of the return pointer pair and seal it too.}\\
    \tmove{\rretc}{\pcreg}\\
    \tcca{\rretc}{5} \text{ //magic number is offset to return code}\\
    \tcseal{\rretc}{\rtmp{1}}\\
    \text{// now clear temporary register and jump to the adversary.}\\
    \tmove{\rtmp{1}}{0}\\
    \txjmp{r_1}{r_2}\\
    \text{// the following is the return code}\\
    \text{// check that the stack pointer is the same we handed out.}\\
    \tgetb{\rtmp{1}}{\rstk}\\
    \tminus{\rtmp{1}}{t_{t1}}{\stkb} \text{ //$\stkb$ is the stack base constant} \\
    \tmove{\rtmp{2}}{\pcreg}\\
    \tcca{\rtmp{2}}{5} \text{ //magic number is the offset to fail} \\
    \tjnz{\rtmp{2}}{\rtmp{1}} \\
    \tcca{\rtmp{2}}{1} \text{ //magic number is the offset to after fail} \\
    \tjmp{\rtmp{2}} \\
    \tfail \\
    \text{// join our stored private stack frame with the rest of the stack (this also finishes the stack pointer check).}\\
    \tsplice{\rstk}{\rstk}{\rdata} \\
    \text{// pop the magic number 42}\\
    \tcca{\rstk}{1}\\
    \text{// clear temporary registers used}\\
    \tmove{\rtmp{2}}{0}\\
    \text{// continue program after invocation.}
\]
\dominique{Return code could be simplified if we had jz in addition to jnz.}
% \dominique{It is a bit annoying that we store the stack end address before invocation if we do not use it.
%   I understand this is for ensuring that our private stack frame is not empty, which is important.
%   But why don't we just push a constant then?
%   Also: perhaps we could only do it if the current private stack frame is empty to begin with?
% \dominique{4-12-2017: update: Lau executed the first suggestion (push a constant instead), and convinced me that the last suggestion didn't make sense (only push if the current private stack is empty), because it is then not clear whether or not to pop in the return code.}
The call code does the following:
\begin{itemize}
\item Store 42 (it could be anything) to the stack (this ensures the stack is non-empty), and decrement the pointer according to convention.
\item Get the current address of the stack pointer and split according to it.
\item Retrieve the seal of the program.
\item cca the seal, so the seal to be used is active 
\item seal our private part of the stack capability.
\item Move the pc out of the pc register, adjust it to point to the first address of the return code, and seal it.
\item clear the temporary register.
\item cross jump to the two specified registers.
\item Upon return:
  \begin{itemize}
  \item get the base of the stack and check that it matches up with the global base of the stack.If not, fail.
  \item splice the returned stack pointer and the stack pointer for our private stack.
  \item adjust stack pointer to first empty address (the address with the end address is considered free)  
  \item clear the temporary registers. (Note that $\rtmp{1}$ is not cleared with 0 as it already contains 0 after the $\stkb$ check. If it didn't contain 0, then the execution would have failed.)

@@

\begin{mathpar}
  \inference{
    \var{comp}_1 = (\mscode[1], \msdata[1], \overline{c_{\mathrm{import},1}}, \overline{c_{\mathrm{export},1}}, \sigrets[1], \sigcloss[1])\\
    \var{comp}_2 = (\mscode[2], \msdata[2], \overline{c_{\mathrm{import},2}}, \overline{c_{\mathrm{export},2}}, \sigrets[2], \sigcloss[2])\\
    \var{comp}_3 = (\mscode[3], \msdata[3], \overline{c_{\mathrm{import},3}}, \overline{c_{\mathrm{export},3}}, \sigrets[3], \sigcloss[3])\\
    \mscode[3] = \mscode[1] \uplus \mscode[2] &
    \msdata[3] = \msdata[1] \uplus \msdata[2] \\
    \overline{c_{\mathrm{import},3}} = (\overline{c_{\mathrm{import},1}} \cup \overline{c_{\mathrm{import},2}}) \setminus \overline{c_{\mathrm{export},3}}&
    \overline{c_{\mathrm{export},3}} = \overline{c_{\mathrm{export},1}} \cup \overline{c_{\mathrm{export},2}}\\
    \sigrets[3] = \sigrets[1] \uplus \sigrets[2] &
    \sigcloss[3] = \sigcloss[1] \uplus \sigcloss[2]
  }{
     \var{comp}_3 = \var{comp}_1 \bowtie \var{comp}_2
  }
\end{mathpar}

\subsection{Programs, contexts, initial execution configuration}
\label{sec:programs}

\begin{definition}[Programs and Contexts]
  We define a program as a $(\var{comp},c_{\mathrm{main}})$ component with an empty import list and $c_{\mathrm{main}}$ is one of the exported symbols.

  Given a component $\var{comp}$, a context is a tuple $(\var{comp'},c_{\mathrm{main}})$ such that $\var{comp'}$ is a component and $(\var{comp} \bowtie \var{comp'},c_{\mathrm{main}})$ is a program.
\end{definition}

@@

\begin{definition}
  \label{def:points-to-call}
  A configuration {\sourcecolor $\Phi = (\ms,\reg,\_,\_)$} is said to point to instruction $\scall{o}{r_1}{r_2}$ in $A$ iff
  \[
    \reg(\pcreg) = ((\_,\_),\_,\eaddr,\aaddr) \tand [\aaddr,\aaddr+\calllen-1] \subseteq A
  \]
and
\[
  \aaddr+\calllen - 1 \leq \eaddr
\]
and
\[
  \callCond{\Phi,r_1,r_2,\aaddr}
\]
\end{definition}
We want to make a conditional full-abstraction statement. The condition should express two things (1) the compilation so far has been reasonable, so we do not have to setup protection against things any reasonable compiler wouldn't do. (2) the compilation so far has taken care of certain checks that we do not have information to place correctly at this point of the compilation.
\begin{definition}
  \label{def:check-stack-addr-before-call}
  A configuration $\src{\Phi}$ is said to check $\stkb$ before calls in $\ta$ for $n$ steps iff\\
  for all $i \leq n$ if
  \[
    \src{\Phi} \nstep[i] \src{\Phi'}
  \]
  and $\src{\Phi'}$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ in $\ta$ for some $\src{r_1}$ and $\src{r_2}$
  % and there exists no $j<i$ s.t.
  % \[
  %   \src{\Phi} \nstep[j] \src{\Phi''}
  % \]
  % where $\src{\Phi''}$ points to $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$ in $A$ for some $\src{r_1'}$ and $\src{r_2'}$,
  then
  \[
    \src{\Phi'}(\src{r_\stk}) = \src{\stkptr{\_,\stkb,\_,\_}}
  \]
  and
  \[
    r_1' \neq \rtmp{1}
  \]
  % and $\src{\Phi'}$ has reasonable calls in $A$ for $n-i$ step.
\end{definition}

\begin{definition}[Use return seals only for call]
  A configuration $\src{\Phi}$ is said to use linker seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
  for all $i \leq n$, if
  \[
    \src{\Phi} \nstep[i] \src{\Phi'}
  \]
  and $\src{\Phi'}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi'(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$, then either ($\src{\Phi'}$ is inside $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$ and $\sigma \in \sigrets$) or $\sigma \in \sigcloss$.
\end{definition}

\begin{definition}[Never pass trusted seals accross the trusted-untrusted boundary]
  A configuration $\src{\Phi}$ is said to use linker seals properly in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
  for all $i \leq n$, if
  \[
    \src{\Phi} \nstep[i] \src{\Phi'} \step \src{\Phi''}
  \]
  and $\src{\Phi'}$ is in $\ta$ and $\src{\Phi''}$ is not in $\ta$ and, then for all $r \in \RegName$
  %\dominique{TODO: $\sigma_b,\sigma_e$ contains a seal that was put in $\Phi$ by the linker.}
  \[
    \src{\Phi}.\reg(r) = \seal{\sigma_b,\sigma_e,\_} \Rightarrow (\sigrets \cup \sigcloss) \mathrel{\#} [\sigma_b,\sigma_e]
  \]
and
%  \dominique{TODO: Also not allowed: passing a capability to a code area that points (a.o.) to a memory location containing a seal like the one above.}
  \[
    \src{\Phi}.\reg(r) = ((\perm,\_),\baddr,\eaddr,\_) \wedge \perm \sqsubseteq \rx \Rightarrow [\baddr,\eaddr] \# \ta
  \]
  and
    \begin{multline*}
      \src{\Phi}.\reg(r) = \sealed{\sigma,\vsc} \wedge \sigma \not\in (\sigrets\cup\sigcloss) \Rightarrow 
      \\  \left(  
      \begin{array}{r l}
        (\vsc = \seal{\sigma_b,\sigma_e,\_} & \Rightarrow (\sigrets\cup\sigcloss) \# [\sigma_b,\sigma_e]) \wedge\\
        ( \vsc = ((\perm,\_),\baddr,\eaddr,\_) \wedge \perm \sqsubseteq \rx & \Rightarrow [\baddr,\eaddr] \# \ta )
      \end{array}
      \right)
  \end{multline*}
\end{definition}

%  \dominique{TODO: Also not allowed: storing a seal or a code capability in memory.}a
\begin{definition}[Never store link seals and code capability]
  A configuration $\src{\Phi}$ is said to not store seals and code capability in $\ta$ with component parameters $\sigrets$, $\sigcloss$ for $n$ steps iff\\
  for all $i \leq n$, if
  \[
    \src{\Phi} \nstep[i] \src{\Phi'}
  \]
  and $\src{\Phi'}$ points to $\src{\tload{r_1}{r_2}}$ in $\ta$, then 
\[
  \Phi'(r_2) = \seal(\sigma_\baddr,\sigma_\eaddr,\_) \Rightarrow (\sigrets\cup\sigcloss) \mathrel{\#} [\sigma_\baddr,\sigma_\eaddr]
\]
and
\[
  \Phi'(r_2) = ((\perm,\_),\baddr,\eaddr,) \wedge \perm \sqsubseteq \rx \Rightarrow [\baddr,\eaddr] \# \ta
\]
\end{definition}

@@

\dominique{19-3-2018: components should contain information about what seals they use and for what (for building return pointers or for building closures). That should facilitate defining reasonable linking (only when seals disjoint), and for defining the condition for full abstraction and the FTLR.}
\begin{definition}
  A program $\program$ with memory segment $\ms_\program$ and $\dom(\ms_\program) = \ta$, is reasonable if for all $n$ and all contexts $\context$ the configuration $\plug{\context}{\program}$ satisfies:
  \begin{itemize}
  \item $\plug{\context}{\program}$ has reasonable calls in $\dom(\ms_\program)$ for $n$ steps.
  \item $\plug{\context}{\program}$ uses linker seals properly for $n$ steps.
  \item $\plug{\context}{\program}$ never passes trusted seals across the boundary.
  \end{itemize}
\end{definition}
